---
title: "Lab 3"
Name: Carmen Chen
Date: 09/10/2021
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## Step 1: Read the data

```{r}
#Install only if we don't have the package
if(!require(data.table)){
  install.packages("data.table")
}

#load required package
library(data.table)

#Only download it if we haven't
if(!file.exists("met_all.gz")){
  
  download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", 
  destfile = "met_all.gz", 
  method="libcurl", 
  timeout = 60
  )

}

met <- data.table::fread("met_all.gz")
```


## Step 2: Checking the dimensions, headers, footers
```{r}
dim(met)
head(met)
tail(met)
```

## Step 3: Take a look at the variables

```{r}
str(met)
```


## Step 4: Take a closer look at the key variables

Looking at categorical variables
```{r}
table(met$year)
table(met$day)
table(met$hour)
```

Looking at continuous variables
```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

Replace elevations with 9999 as NA
```{r}
met[met$elev==9999.0] <- NA
#met[elev == 9999.0, elev := NA]
summary(met$elev)
```

The highest weather station is at `r max(met$elev, na.rm=TRUE)` `r ifelse(runif(1) > .5, "ft", "mts")`.

Now we need to fix the temperature.
```{r}
met <- met[temp>-40][order(temp)]
summary(met[, .(lat, lon, wind.sp, temp, elev)])
head(met)
```

##Step 5: Check the data against an external data source

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```


##Step 6: Calculate summary statistics

```{r}
elev <- met[elev==max(elev)]
summary(elev)
```

```{r}
met[elev == max(elev, na.rm=TRUE), .(
  temp_wind = cor(temp, wind.sp, use="complete"),  
  temp_day = cor(temp, day, use="complete"),
  temp_hour = cor(temp, hour, use="complete"),
  wind_day = cor(wind.sp, day, use="complete"),
  wind_hour = cor(wind.sp, hour, use="complete")
)]
```

## Step 7: Exploratory graphs

```{r}
hist(met$elev, breaks=100)
```



```{r}
hist(met$temp)
```

```{r}
hist(met$wind.sp)
```

Look at where the weather station with highest elevation is located.

```{r}
library(leaflet)
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

