---
title: "lab05"
author: "Carmen Chen"
date: "9/24/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Read in the data

```{r load the packages}
library(data.table)
library(dtplyr)
library(dplyr)
```

```{r stations-data, cache=TRUE}
#met
if (!file.exists("../met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "../met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../met_all.gz")

# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r merge the data}
met <- merge(
  x = met,
  y = stations,
  all.x = TRUE, all.y = FALSE,
  by.x = "USAFID", by.y = "USAF"
)
met %>% nrow()
```

## Question 1: Representative station for the US

What is the median station in terms of temperature, wind speed, and atmospheric pressure? Look for the three weather stations that best represent continental US using the quantile() function. Do these three coincide?


First, generate a representative version of each station. We will use the averages (median could also be a good way to represent it, but it will depend on the case)

```{r collapsing-by-station}
station_averages <- met[, .(
  temp = mean(temp, na.rm = TRUE),
  wind.sp = mean(wind.sp, na.rm = TRUE),
  atm.press = mean(atm.press, na.rm = TRUE),
  lon = mean(lon, na.rm = TRUE),
  lat = mean(lat, na.rm = TRUE)
), by = .(USAFID)]
station_averages
```

Now, we need to identify the quantiles per variable

```{r quantiles}
medians <- station_averages[, .(
  temp_50 = quantile(temp, probs = .5, na.rm = TRUE),
  wind.sp_50 = quantile(wind.sp, probs = .5, na.rm = TRUE),
  atm.press_50 = quantile(atm.press, probs = .5, na.rm = TRUE)
)]

medians
```
Now we can find the stations that are the closest to these. (hint: `which.min()`)

```{r median-temp}
station_averages[, temp_dist := abs(temp - medians$temp_50)]
median_temp_station <- station_averages[order(temp_dist)][1]
median_temp_station
```
```{r median-wind.sp}
station_averages[, wind.sp_dist := abs(wind.sp - medians$wind.sp_50)]
median_wind.sp_station <- station_averages[order(wind.sp_dist)][1]
median_wind.sp_station
```

```{r median-atm.press}
station_averages[, atm.press_dist := abs(atm.press - medians$atm.press_50)]
median_atm.press_station <- station_averages[order(atm.press_dist)][1]
median_atm.press_station
```

The median temperature station is `r median_temp_station$USAFID`.
The median wind speed station is `r median_wind.sp_station$USAFID`.
The median atm.press station is `r median_atm.press_station$USAFID`.



## Question 2: Representative station per state

Just like the previous question, you are asked to identify what is the most representative, the median, station per state. This time, instead of looking at one variable at a time, look at the euclidean distance. If multiple stations show in the median, select the one located at the lowest latitude.


We first need to recover the state variable, by MERGING.

```{r}
station_averages <- merge(x = station_averages, y = stations, by.x = "USAFID", by.y = "USAF", all.x = FALSE, all.y = TRUE)
station_averages
```

```{r}
station_averages[, temp_50 := quantile(temp, probs = .5, na.rm = TRUE), by = STATE ]
station_averages[, wind.sp_50 := quantile(wind.sp, probs = .5, na.rm = TRUE), by = STATE ]
station_averages[, atm.press_50 := quantile(atm.press, probs = .5, na.rm = TRUE), by = STATE ]
```

Now, the euclidean distance... $\sqrt{\sum_i(x_i - y_i)^2}$

```{r}
station_averages[, eudist := sqrt(
  (temp - temp_50)^2 + (wind.sp - wind.sp_50)^2 +(atm.press - atm.press_50)^2
)]
representative <- setDT(station_averages)[, .SD[which.min(eudist)], by=STATE]
knitr::kable(representative)
```


## Question 3: In the middle?
For each state, identify what is the station that is closest to the mid-point of the state. Combining these with the stations you identified in the previous question, use leaflet() to visualize all ~100 points in the same figure, applying different colors for those identified in this question.

```{r midpoint}
station_averages[, lat_50 := quantile(lat, probs = .5, na.rm = TRUE), by = STATE]
station_averages[, lon_50 := quantile(lon, probs = .5, na.rm = TRUE), by = STATE]

```

```{r euclidean distance}
station_averages[, latlon_eudist := sqrt(
  (lat - lat_50)^2 + (lon - lon_50)^2
)]
midpoint <- setDT(station_averages)[, .SD[which.min(latlon_eudist)], by=STATE]
knitr::kable(midpoint)
```

```{r combine}
representative[, label := "Wind speed, temperature, and atmospheric pressure"]
midpoint[, label := "Midpoint"]
combine <- rbind(representative, midpoint, fill = TRUE)
```

```{r map}
library(leaflet)
leaflet(combine) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircles(
    lat = ~lat, lng = ~lon, color = ~ifelse(label == "Midpoint", 'blue', 'green'))

```


## Question 4: Means of means
Using the quantile() function, generate a summary table that shows the number of states included, average temperature, wind-speed, and atmospheric pressure by the variable “average temperature level,” which you’ll need to create.

Start by computing the states’ average temperature. Use that measurement to classify them according to the following criteria:

low: temp < 20
Mid: temp >= 20 and temp < 25
High: temp >= 25
Once you are done with that, you can compute the following:

Number of entries (records),
Number of NA entries,
Number of stations,
Number of states included, and
Mean temperature, wind-speed, and atmospheric pressure.


```{r}
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]
met[, temp_cat := fifelse(
  state_temp < 20, "low-temp", 
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]
```


Let's make sure that we don't have NAs.

```{r}
table(met$temp_cat, useNA = "always")
```


Now, let's summarize

```{r}
table <- met[, .(
  N_entries = .N,
  N_NA = sum(is.na(temp)),
  N_stations = length(unique(USAFID)),
  N_sates = length(unique(STATE)),
  Mean_temp = mean(temp, na.rm = TRUE),
  Mean_wind.sp = mean(wind.sp, na.rm = TRUE),
  Mean_atm.press = mean(atm.press, na.rm = TRUE)
  ), by = temp_cat]
knitr::kable(table)
```


