---
title: "LAB07"
author: "Carmen Chen"
date: "10/8/2021"
output: 
  github_document: default
  html_document:
    html_preview: FALSE
  word_document: default
always_allow_html: TRUE
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
library(tidyverse)
library(tidytext)
```

if(knitr::is_html_output(excludes = "gfm)){

}



#Question 1: How many sars-cov-2 papers?
Build an automatic counter of sars-cov-2 papers using PubMed. You will need to apply XPath as we did during the lecture to extract the number of results returned by PubMed in the following web address:

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/section[1]/div[2]/div[1]/span") #inspect -> copy full x path

# Turning it into text
#or use xml2::xml_text(counts)
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
stringr::str_extract(counts, "[[:digit:],]+")
stringr::str_replace(counts, "[^[:digit:]]+([[:digit:]]+),([[:digit:]]+)[^[:digit:]]+", "\\1\\2")

```

#Question 2: Academic publications on COVID19 and Hawaii

```{r hawaii}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db = "pubmed",
    term = "covid19 hawaii",
    retmax = 1000)
)

GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/",
  path = "entrez/eutils/esearch.fcgi",
  query = list(
    db = "pubmed",
    term = "covid19 hawaii",
    retmax = 1000)
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```


#Question 3: Get details about the articles
The Ids are wrapped around text in the following way: <Id>... id number ...</Id>. we can use a regular expression that extract that information. Fill out the following lines of code:

```{r eval=FALSE}
ids_list <- xml2::as_list(ids) 
ids_char <- as.character(ids)
```


```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
#stringr::str_remove_all(ids, "</?Id>")
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
head(ids)
paste(ids, collapse =",")
```

```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
   db = "pubmed",
   id =  I(paste(ids, collapse = ",")),
   retmax = 1000,
   rettype = "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

#Question 4: Distribution of universities, schools, and departments

```{r univ-insti}
institution <- str_extract_all(
  str_to_lower(publications_txt),
  "university\\s+of\\s+(southern|new|northern|the)?\\s*[[:alpha:]-]+|[[:alpha:]-]+\\s+institute\\s+of\\s+[[:alpha:]-]+"
  ) 
institution <- unlist(institution)
table(institution)
```

```{r school-dep}
schools_and_deps <- str_extract_all(
  str_to_lower(publications_txt),
  "school\\s+of\\s+(southern|new|northern|the)?\\s*[[:alpha:]-]+|[[:alpha:]-]+\\s+department\\s+of\\s+[[:alpha:]-]+"
  )
table(schools_and_deps)
```

#Question 5: Form a database

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") # '</?[[:alnum:]- ="]+>'
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")
```

```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
```

```{r}
database <- data.frame(
  PubMed = ids,
  Title = titles,
  Abstract = abstracts
)
knitr::kable(database[1:20,], caption = "Some papers about Covid19 and Hawaii")
```

Rscript --vanilla -e 'rmarkdown::render("index.Rmd", output_format = "all")'
