---
title: "Lab 06"
author: "Carmen Chen"
date: "10/1/2021"
output: github_document
always_allow_html: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com")) 
library(tidyverse)
library(tidytext)
library(tibble)
library(dplyr)
```

##Download the data
```{r get-data, cache=TRUE}
fn <- "mtsamples.csv"
if(!file.exists(fn))
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv", destfile = fn)

mtsamples <- read.csv(fn) #read data from a data frame
mtsamples <- as_tibble(mtsamples)
```


#Question 1: What specialties do we have?
We can use count() from dplyr to figure out how many different catagories do we have? Are these catagories related? overlapping? evenly distributed?


```{r dist-of-specialities}
specialties <- mtsamples %>%
  count(medical_specialty)

specialties %>%
  arrange(desc(n)) %>%
  top_n(15) %>%
  knitr::kable()
```

There are `r nrow(specialties)` specialties. Let's take a look at the distribution:

```{r dist1}
#Method 1 (not that pretty)
ggplot(mtsamples, aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip() #for better reading the categories
```

```{r dist2}
#Method 2
ggplot(specialties, aes(x = n, y = fct_reorder(medical_specialty, n))) + #x is frequency, y is the label, sorting the medical_specialty according to n
  geom_col() #plot column
```

These are not evenly distributed.


#Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words

```{r token-transcript}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word, n))) +
  geom_col()

```

The word "patient" seems to be important, but we observe a lot of stop words.

#Question 3
Redo visualization but remove stopwords before
Bonus points if you remove numbers as well
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

```{r}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  top_n(20) %>%
  ggplot(aes(x = n, y = fct_reorder(word, n))) +
  geom_col()
  
```

Looking better, but we don't like the numbers

```{r}

```























How to push the file in Terminal using command lines:

git add lab06/README
git status
git commit -a -m "Starting with lab06"
git push


How to remove the file on track:

git rm --cache -r lab06/README_cache


